---
title: "Examples"
---

This page provides an example of how to compute and visualize distinction centrality using the classic Medici family network from Padgett & Ansell (1993).

```{r}
#| message: false
#| warning: false

library(igraph)
library(dplyr)
library(readr)
library(patchwork)
source("Functions/distinction.R")
source("Functions/plot.graph.norm.R")

# Load the Medici dataset
medici_edges <- read_csv("Data/MediciEdgeList.csv")
g <- graph_from_data_frame(medici_edges, directed = FALSE)

# Calculate distinction centrality for the full network
results_full <- distinction(g)
```

### Top Families (Full Network)

```{r}
#| echo: false
results_full %>%
  arrange(desc(scd)) %>%
  head(5) %>%
  knitr::kable()
```

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 6
#| out-width: "100%"
set.seed(42)
plot.graph.norm(g, ts = 4, dist_data = results_full) + ggplot2::ggtitle("Full Florentine Network")
```

### Top Families (Network without Medici)

```{r}
#| message: false
#| warning: false

# Load and analyze the network without the Medici
nomedici_edges <- read_csv("Data/NoMediciEdgeList.csv")
g_no <- graph_from_data_frame(nomedici_edges, directed = FALSE)
results_no <- distinction(g_no)
```

```{r}
#| echo: false
results_no %>%
  arrange(desc(scd)) %>%
  head(5) %>%
  knitr::kable()
```

```{r}
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 6
#| out-width: "100%"
set.seed(42)
plot.graph.norm(g_no, ts = 4, dist_data = results_no) + ggplot2::ggtitle("Florentine Network (Medici Removed)")
```

### Analysis
The comparison highlights how the network hierarchy shifts in the absence of the Medici. In the full network, the Medici command the highest distinction score (shown in red in the plot), acting as the central broker. When removed, the distinction scores of the remaining elite families are recalculated based on their new relative network positions. Specifically, families like the Albizzi and Guasconi experience significant gains in distinction centrality, taking over as the new brokers in the fragmented structure. This illustrates how individual distinction is intrinsically linked to the presence or absence of key central players.
